---
title: "brennas_OriginalHomeworkCode_04"
author: "brennastallings"
date: "10/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is for [Homework 4](https://fuzzyatelin.github.io/bioanth-stats/homework-04.html)

>Question 1

#Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data, using the following guidelines:
~There is a sample large function in [Module 11](https://fuzzyatelin.github.io/bioanth-stats/module-11/module-11.html)

Your function should take the following arguments: p1 and n1 (no default) representing the estimated proportion and sample size (i.e., based on your sample data); p2 and n2 (both defaulting to NULL) that contain a second sample’s proportion and sample size data in the event of a two-sample test; p0 (no default) as the expected value for the population proportion; and alternative (default “two.sided”) and conf.level (default 0.95), to be used in the same way as in the function t.test().
This sets up the beginning, skeleton, function
```{r}
z.prop.test <- function(p1, n1, p2= NULL, n2= NULL, p0, alternative="two.tailed", conf.level=0.95){
 
                }
```

I wrote this code piece by piece, so these are so I can test individual equations
```{r}
p1 = 0.8
p2 = 0.3
p0 = 0.4
n1 = 25
n2 = 5
conf.level = 0.95
```

When conducting a two-sample test, it should be p1 that is tested as being smaller or larger than p2 when alternative=“less” or alternative=“greater”, the same as in the use of x and y in the function t.test().

```{r echo=F}
#  if (alternative == "two.tailed")
#  {
   # CODE FOR 2 TAILED
#  }
#  if (alternative == "greater"| "less")
#  {
#    pstar<- (sum(p1*n1)+sum(p2*n2))/(n1+n2)
#    #sum(p1*n1) gives the number of sucesses, which is x1
#    #n1 is the number of observations
#    z<- (p2 - p1)/sqrt((pstar * (1-pstar)) * (1/length(n1) + 1/length(n2)))
#    p<- 1-pnorm(z, lower.tail = T) + pnorm(z,lower.tail = F)
#    lower <- ((p2-p1)-qnorm(conf.level) * sqrt((p2-p1)*(1-(p2-p1)))/n1)
#    upper <- ((p2-p1)+qnorm(conf.level) * sqrt((p2-p1)*(1-(p2-p1)))/n1)
#    ci <- c(lower, upper)
#  }

#the math in here gives a real number, so that works at least
```

The function should perform a one-sample Z-test using p1, n1, and p0 if either p2 or n2 (or both) is NULL.

```{r echo=F}
   if (is.null(n2)|is.null(p2))
  {
    z<- (p1-p0)/sqrt((p0*(1-p0))/n1)
    p<- pnorm(z, lower.tail = T)
    lower <- (p1-qnorm(conf.level) * sqrt(p1*(1-p1))/n1)
    upper <- (p1+qnorm(conf.level) * sqrt(p1*(1-p1))/n1)
    ci <- c(lower, upper)
  }
```

The function should contain a check for the rules of thumb we have talked about (n∗p>5 and n∗(1−p)>5) to ensure the validity of assuming the normal distribution in both the one- and two-sample settings. If this is violated, the function should still complete but it should also print an appropriate warning message.
#this should be if statements
```{r echo=F}
if (n1*p0 < 5)
  {
  print("Caution, n*p in less than 5, violating normality assumption")
  }
if (n2*p0 < 5)
  {
  print("Caution, n*p in less than 5, violating normality assumption")
  }
if (n1*(1-p0) < 5)
  {
  print("Caution, n*(1-p) in less than 5, violating normality assumption")
}
if (n2*(1-p0) < 5)
  {
  print("Caution, n*(1-p) in less than 5, violating normality assumption")
}

```

The function should return a list containing the members Z (the test statistic), P (the appropriate p value), and CI (the two-sided CI with respect to “conf.level” around p1 in the case of a one-sample test and around p2-p1 in the case of a two-sample test). For all test alternatives (“two.sided”, “greater”, “less”), calculate symmetric CIs based on quantiles of the normal distribution rather than worrying about calculating single-limit confidence bounds.

#this can be put after all if statemnets
```{r echo=F}
#values<- c("z-stat", z, "p value", p, "conf. intervals", ci)
#print(values)
```

What i want for final function to be arranged:
<br></br>
1.if alt= less, calculate values
<br></br>
2.if alt= greater, calculate values
<br></br>
3.if alt= two.tailed, calculate values
<br></br>
4.if p2 or n2 is null, calulate values
<br></br>
5.if there are errors, print warning
<br></br>
6.print values

This is for my final function. I am trying to add it one piece at a time
```{r}
z.prop.test <- function(p1, n1, p0, p2= NULL, n2= NULL, alternative="two.tailed", conf.level=0.95){
   if (alternative == "two.tailed")
  {
    z<- (p1-p0)/sqrt((p0*(1-p0))/n1)
    p<- pnorm(z, lower.tail = T)
    lower <- (p1-qnorm(conf.level) * sqrt(p1*(1-p1))/n1)
    upper <- (p1+qnorm(conf.level) * sqrt(p1*(1-p1))/n1)
    ci <- c(lower, upper)
   }
#this is for if n2 or p2 is null
  if (is.null(n2)|is.null(p2))
  {
    z<- (p1-p0)/sqrt((p0*(1-p0))/n1)
    p<- pnorm(z, lower.tail = T)
    lower <- (p1-qnorm(conf.level) * sqrt(p1*(1-p1))/n1)
    upper <- (p1+qnorm(conf.level) * sqrt(p1*(1-p1))/n1)
    ci <- c(lower, upper)
  }
#this is for if p1 is less than p2
    if (alternative == "less")
  {
    pstar<- (sum(p1*n1)+sum(p2*n2))/(n1+n2)
    #sum(p1*n1) gives the number of sucesses, which is x1
    #n1 is the number of observations
    z<- (p2 - p1)/sqrt((pstar * (1-pstar)) * (1/length(n1) + 1/length(n2)))
    p<- 1-pnorm(z, lower.tail = T) + pnorm(z,lower.tail = F)
    lower <- ((p2-p1)-qnorm(conf.level) * sqrt((p2-p1)*(1-(p2-p1)))/n1)
    upper <- ((p2-p1)+qnorm(conf.level) * sqrt((p2-p1)*(1-(p2-p1)))/n1)
    ci <- c(lower, upper)
  }
 #this is for if p1 is greater than p2
   if (alternative == "greater")
  {
    pstar<- (sum(p1*n1)+sum(p2*n2))/(n1+n2)
    #sum(p1*n1) gives the number of sucesses, which is x1
    #n1 is the number of observations
    z<- (p1 - p2)/sqrt((pstar * (1-pstar)) * (1/length(n1) + 1/length(n2)))
    p<- 1-pnorm(z, lower.tail = T) + pnorm(z,lower.tail = F)
    lower <- ((p1-p2)-qnorm(conf.level) * sqrt((p1-p2)*(1-(p1-p2)))/n1)
    upper <- ((p1-p2)+qnorm(conf.level) * sqrt((p1-p2)*(1-(p1-p2)))/n1)
    ci <- c(lower, upper)
  }
values<- c("z-stat", z, "p value", p, "conf. intervals", ci)
print(values)

#These are my error messages
if (n1*p0 < 5)
  {
  print("Caution, n*p in less than 5, violating normality assumption")
  }
if (n1*(1-p0) < 5)
  {
  print("Caution, n*(1-p) in less than 5, violating normality assumption")
  }
#these only need to show up if using p2 and n2
if (!is.null(n2))
{
  if (n2*p0 < 5)
  {
  print("Caution, n*p in less than 5, violating normality assumption")
  }
  if (n2*(1-p0) < 5)
  {
  print("Caution, n*(1-p) in less than 5, violating normality assumption")
  }
}
}
```
I think this worked!!!
Math may be wrong though

>Question 2

#The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species’ brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both longevity~brain size and log(longevity)~log(brain size):

First we load any necessary libraries and the data set
```{r}
library(ggplot2)
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall19/KamilarAndCooperData.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).
```{r}
#Fitting the regression model
#First Model
m<- lm(MaxLongevity_m ~ Brain_Size_Species_Mean, data= d)


#Graphing
g <- ggplot(data = d, aes(x = MaxLongevity_m, y = Brain_Size_Species_Mean ))
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g <- g + geom_text(aes(x = 0, y = 500), hjust = 0,label =  "Longevity ~ Brain Size" )
g

#log MOdel
n<- lm(log(MaxLongevity_m) ~ log(Brain_Size_Species_Mean), data= d)
h <- ggplot(data = d, aes(x = log(MaxLongevity_m), y = log(Brain_Size_Species_Mean) ))
h <- h + geom_point()
h <- h + geom_smooth(method = "lm", formula = y ~ x)
h <- h + geom_text(aes(x = 0, y = 7.5), hjust = 0,label =  "log(Longevity) ~ log(Brain Size)" )
h

#want to organize the graphs side by side
par(mfrow = c(1,2))

```

Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.
```{r echo=F}
m
```
β1=1.218
This means that for every month of longevity, the brain size increases 1.218g
```{r echo=F}
t <- coef(summary(m))
t <- data.frame(unlist(t))
colnames(t) <- c("Est", "SE", "t", "p")
t
```
Looking at the t-test values, the outcome would be to reject H0.
```{r echo=F}
ci <- confint(m, level = 0.90)  # using the results of lm()
ci
```

#These are for the log funcion
```{r echo=F}
n
```
β1=0.2341
This means that for the log of every unit of longevity, the log of the brain size gose up 0.2341 units
```{r echo=F}
t <- coef(summary(n))
t <- data.frame(unlist(t))
colnames(t) <- c("Est", "SE", "t", "p")
t
```
Looking at the t-test values, the outcome would be to reject H0.
```{r echo=F}
ci <- confint(n, level = 0.90)  # using the results of lm()
ci
```


Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.
```{r}
#first equation
#this creates the predictions for the CI Intervals
ci <- predict(m, newdata = data.frame(Brain_Size_Species_Mean = d$Brain_Size_Species_Mean), interval = "confidence", level = 0.90)  # for a vector of values
h_hat <- predict(m, newdata = data.frame(Brain_Size_Species_Mean = d$Brain_Size_Species_Mean))
#this puts it into a df and organizes it
df <- data.frame(cbind(d$Brain_Size_Species_Mean, d$MaxLongevity_m, h_hat))
names(df) <- c("x", "y", "yhat")
df <- cbind(df, ci)
names(df) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr")
#this creates the predictions for the PI INtervals
pi <- predict(m, newdata = data.frame(Brain_Size_Species_Mean = d$Brain_Size_Species_Mean), interval = "prediction", 
    level = 0.90)  # for a vector of values
pf <- cbind(df, pi)
names(pf) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr", "PIfit", "PIlwr", 
    "PIupr")
#This starts the plotting
g <- ggplot(data = df, aes(x = x, y = y))
g <- g + geom_point(alpha = 1/2)
g <- g + geom_line(aes(x = x, y = CIfit), colour = "black")
g <- g + geom_line(aes(x = x, y = CIlwr), colour = "blue")
g <- g + geom_line(aes(x = x, y = CIupr), colour = "blue")
g <- g + geom_line(data = pf, aes(x = x, y = PIlwr), colour = "red")
g <- g + geom_line(data = pf, aes(x = x, y = PIupr), colour = "red")
g <- g + scale_colour_manual("", 
                      breaks = c("Confidence Interval", "Prediction Interval"),
                      values = c("blue", "red"))
#This line doesnt work-YET
g <- g + theme (
    legend.position = c(0.95, 0.95),
    legend.justification = c("right", "top"))
g
```
# cannot make the legend show up

```{r}
#log equation
cil <- predict(n, newdata = data.frame(Brain_Size_Species_Mean = log(d$Brain_Size_Species_Mean)), interval = "confidence", level = 0.90)

hl_hat <- predict(n, newdata = data.frame(Brain_Size_Species_Mean = log(d$Brain_Size_Species_Mean)))

dfl <- data.frame(cbind(log(d$Brain_Size_Species_Mean), log(d$MaxLongevity_m), hl_hat))
names(dfl) <- c("x", "y", "yhat")

dfl <- cbind(dfl, cil)

names(dfl) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr")
#this creates the predictions for the PI INtervals
pil <- predict(n, newdata = data.frame(Brain_Size_Species_Mean = log(d$Brain_Size_Species_Mean)), interval = "prediction", 
    level = 0.90) 
pfl <- cbind(dfl, pil)

names(pfl) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr", "PIfit", "PIlwr", 
    "PIupr")
#this is the actual graphing
g <- ggplot(data = dfl, aes(x = x, y = y))
g <- g + geom_point(alpha = 1/2)
g <- g + geom_line(aes(x = x, y = CIfit), colour = "black")
g <- g + geom_line(aes(x = x, y = CIlwr), colour = "blue")
g <- g + geom_line(aes(x = x, y = CIupr), colour = "blue")
g <- g + geom_line(data = pfl, aes(x = x, y = PIlwr), colour = "red")
g <- g + geom_line(data = pfl, aes(x = x, y = PIupr), colour = "red")
g <- g + scale_colour_manual("", 
                      breaks = c("Confidence Interval", "Prediction Interval"),
                      values = c("blue", "red"))
g

```
#these lines are wrong, they dont look like they fit
#the legend continues to not work

Produce a point estimate and associated 90 percent PI for the longevity of a species whose brain weight is 800 gm. Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?
```{r echo=F}
ci <- predict(m, newdata = data.frame(Brain_Size_Species_Mean = 800), interval = "confidence", 
    level = 0.90)  # for a single value
ci
```
I dont trust this model to predict accuratly, as the CI interval are very large, since the data doesnt have much information that far out. The CI is more than 100 months above or below the point estimate, which is 11 years.

For the log function
```{r echo=F}
#for log function
ci <- predict(n, newdata = data.frame( Brain_Size_Species_Mean = log(800)), interval = "confidence",
    level = 0.90)  # for a single value
ci
```
I wouldnt trust this becasue I cant understand it. If you log transform them back...
```{r}
exp(5.32379)
exp(5.259948)
exp(5.387632)
```

Looking at your two models, which do you think is better? Why?
I prefer the first model, as it is easier to interpret. It may be becasue my math is wrong, but the CIs for the log function seem large in the graphs.

>5 Challenges

<br></br>
1. I had serious issuse with the large function, I had trouble with the statements for if n2 or p2 were null. I think it works now but I am unsure.
<br></br>
2. I am still having trouble getting the legend onto the graphs for the CI and PIs
<br></br>
3. My graphed intervals do not match the data at all for the log function and Im not sure where my problem is.
<br></br>
4. I kept running into issues when trying to create the dataframe with all the CI and PI calculations. I had an issue with the various names for objects that I have fixed
<br></br>
5. Im still a bit confused about the function and Im not sure if it calculates stats properly.



```{r echo = F}
#this is just the sample from class
bro <- function(x){
  if(x == "bro"){
    "bro"
  }
  else{
    "blank stare"
  }
}
```

